---
title: "Distance to coastline"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("dist2coast.R")
```

## Description

This function is a wrapper for a few other functions and calculates the distance of geo points to the nearest coastline. It is meant to be used for marine data. In fact it calculates the distance to the nearest line. When using the `coastline = "mapdata"` method, for points on land the result may well be the distance to a national border that is not the coastline! The method `coastline = "ne"` uses lines representing only the coastline, but with much lower resolution. 

The input (and the coastline) will be regarded as WGS84. This can be reprojected to the planar UTM 32 references system by choosing `as_UTM32 = TRUE`, which will make the processing faster, especially when global coastline is not filtered via `coastline_crop`, but it is only recommended for smaller spatial extend, or for a quick overview.

The result is either a vector of distances (in m) corresponding to the input coordinates (default) or a matrix of distances between each point and each line of the coastline data (this may actually be quite useless because the coastline data is not output). A figure of the results can be plotted with the `plot = TRUE` argument.

NOTE: requires packages `mapdata` (for `coastline = "mapdata"`) and `rnaturalearthdata` (for `coastline = "ne"`) to be installed.

## Arguments

*lons* -- vector of longitudes

*lats* -- vector of latitudes

*coastline_crop = NULL* -- numeric vector with xmin, ymin, xmax and ymax for cropping the coastline to a bounding box, e.g. c(xmin = -1, ymin = 50, xmax = 11, ymax = 60), elements do not have to be named but must be in the correct order

*coastline = "ne"* -- coastline source; "ne" for Natural Earth (www.naturalearthdata.com) (faster), "mapdata" for using `map('world')` (more precise)

*as_utm32 = FALSE* -- transform both points and coast to UTM32, i.e. a planar projection, for more speed

*output = "mindist"* -- "mindist" for distance to nearest coastline, "distmat" for matrix of distances between points (in rows) and every line (in cols)

*plot = FALSE* -- should the result be shown in a plot

## Value

When `output = "mindist"` a vector of distances in m to the nearest coastline corresponding to the input points.

When `output = "distmat"` a matrix of distances in m with the rows corresponding to input points and columns corresponding to each line element of the coastline data.

When `plot = TRUE` a plot will be output in the graphic device.

## Examples

Dummy points:
```{r}
lons <- seq(2, 9, length.out = 10)
lats <- seq(52, 58, length.out = 10)
```

Default result
```{r}
dist <- dist2coast(lons, lats)
dist
```

Result with finer coastline, filtered coastline, and plot output
```{r}
dist <- dist2coast(lons, lats, coastline_crop = c(-5, 50, 15, 60), coastline = "mapdata", plot = TRUE)
dist
```

